Revealing Benefits, Costs, and Tradeoffs of Spatial Management in Marine Systems Arrising from Coupled Economic and Ecological Dynamics

### 2023-01-10 03_marlin_paper.qmd
out for internal

---
title: "Revealing Benefits, Costs, and Tradeoffs of Spatial Management in Marine Systems Arrising from Coupled Economic and Ecological Dynamics"
author:
    - Daniel Ovando
    - Darcy Bradley
    - Echelle Burns
    - Lennon Thomas
    - James Thorson
date: "Updated `r Sys.Date()`"
format:
  pdf:
    toc: true
    fig-pos: 'H'
  html: default
  docx: default
execute:
  echo: false
  message: false
  warning: false
  cache: false
bibliography: references.bib
params:
  run_name: ["v0.5"]
abstract: "Abstract: Designing effective spatial management strategies is challenging because marine ecosystems are highly dynamic and opaque, and extractive entities such as fishing fleets respond endogenously to ecosystem changes in ways that depend upon ecological and policy context. We present a tool, `marlin`, that can be used to efficiently simulate the bio-economic dynamics of marine systems in support of both management and research. We demonstrate `marlin’s` capabilities by focusing on two case studies on the conservation and food production impacts of marine protected areas (MPAs): a coastal coral reef and a pelagic tuna fishery. In the coastal coral reef example, we explore how heterogeneity in species distributions and fleet preferences can affect distributional outcomes of MPAs. In the pelagic case study, we show how our model can be used to assess the climate resilience of different MPA design strategies, as well as the climate sensitivity of different fishing fleets. This paper demonstrates how intermediate complexity simulation of coupled bio-economic dynamics can help communities predict and potentially manage trade-offs between conservation, fisheries yields, and distributional outcomes of management policies affected by spatial bio-economic dynamics."
header-includes:
  - \usepackage{setspace}\doublespacing
  - \usepackage{lineno}\linenumbers
---

@@

@@

While users can adapt the general equations and ideas presented here as
they see fit, the software package accompanying this paper,
[`marlin`](https://danovando.github.io/marlin/), is written in R
[@rcoreteam2021] and C++, integrated through the `Rcpp` package
[@eddelbuettel2018].

@@

```{r}
#| label: fig-coral-results
#| fig-cap: "Percent change in yield per fleet (A) and depletion per species (B) as a function of MPA size and MPA design strategy for the coastal coral reef case study. Percent change in total yield across both fleets (y-axis) and total depletion across all species (x-axis) by MPA size (point size) and placement strategy (point color) (C)." 
#| fig-height: 4
#| fig-width: 6

@@

coral_frontier <- coral_results %>% 
  group_by(prop_mpa, placement_strategy) %>% 
  summarise(biodiv = sum(biodiv), yield = sum(yield)) %>% 
  group_by(placement_strategy) %>% 
  mutate(delta_yield = yield / yield[prop_mpa == 0] - 1,
         delta_biodiv = biodiv / biodiv[prop_mpa == 0] - 1) %>% 
  ungroup() %>% 
    mutate(placement_strategy= fct_relabel(placement_strategy, titler)) %>% 
  ggplot(aes(delta_biodiv, delta_yield, fill = placement_strategy, size = prop_mpa)) + 
  geom_hline(aes(yintercept = 0), linetype = 2) +
  geom_point(alpha = 0.8, shape = 21) + 
  scale_x_continuous(name = "Change in Total Depletion", labels = scales::label_percent(accuracy = 1), guide = guide_axis(n.dodge = 2)) +
  scale_y_continuous(name = "Change in Yield", labels = scales::label_percent(accuracy = 1)) + 
  scale_fill_viridis_d(name = '')+
  scale_size(name = "MPA Size", labels = scales::label_percent(accuracy = 1), breaks = seq(0.25,1, by = 0.25)) +
  guides(fill = guide_legend(order = 1, override.aes = list(size = 4)),
         size = guide_legend(order = 2, ncol = 2, byrow = TRUE)) + 
  labs(tag = "C")  + 
  theme(
    legend.position = c(0.725, .85),
    legend.background = element_rect(fill = "transparent", color = "black"),
    legend.text = element_text(size = 8)
  )

@@

```{r}
#| label: fig-blue-results
#| fig-cap: "Change in yield (A) and spawning biomass (B) as a function of MPA size and design strategy by fishing fleet and species. Blue lines indicate impacts of MPAs under status qup habitat, red impacts under climate-driven range shift."
#| fig-height: 6
#| fig-width: 6

>>>>

---
title: "Revealing Benefits, Costs, and Tradeoffs of Spatial Management in Marine Systems Arrising from Coupled Economic and Ecological Dynamics"
author:
    - Daniel Ovando
    - Darcy Bradley
    - Echelle Burns
    - Lennon Thomas
    - James Thorson
date: "Updated `r Sys.Date()`"
format:
  pdf:
    toc: true
    fig-pos: 'H'
  html: default
  docx: default
execute:
  echo: false
  message: false
  warning: false
  cache: false
bibliography: references.bib
params:
  run_name: ["v1.0"]
abstract: "Abstract: Designing effective spatial management strategies is challenging because marine ecosystems are highly dynamic and opaque, and extractive entities such as fishing fleets respond endogenously to ecosystem changes in ways that depend upon ecological and policy context. We present a tool, `marlin`, that can be used to efficiently simulate the bio-economic dynamics of marine systems in support of both management and research. We demonstrate `marlin’s` capabilities by focusing on two case studies on the conservation and food production impacts of marine protected areas (MPAs): a coastal coral reef and a pelagic tuna fishery. In the coastal coral reef example, we explore how heterogeneity in species distributions and fleet preferences can affect distributional outcomes of MPAs. In the pelagic case study, we show how our model can be used to assess the climate resilience of different MPA design strategies, as well as the climate sensitivity of different fishing fleets. This paper demonstrates how intermediate complexity simulation of coupled bio-economic dynamics can help communities predict and potentially manage trade-offs between conservation, fisheries yields, and distributional outcomes of management policies affected by spatial bio-economic dynamics."
header-includes:
  - \usepackage{setspace}\doublespacing
  - \usepackage{lineno}\linenumbers
---

@@

The front end of the software package accompanying this paper,
[`marlin`](https://danovando.github.io/marlin/), is written in R
[@rcoreteam2021] to facilitate use, while the underlying population model is written in C++, integrated through the `Rcpp` package
[@eddelbuettel2018], for speed. Users on standard computers should be able to simulate one age-structured population distributed across a 100 patch system for 20 years in fractions of a second.

@@

@@

```{r}
#| label: fig-coral-results
#| fig-cap: "Percent change in yield per fleet (A) and depletion (spawning biomass divided by unfished spawning biomass) per species (B) as a function of MPA size and MPA design strategy for the coastal coral reef case study. Percent change in total yield across both fleets (y-axis) and total depletion across all species (x-axis) and placement strategies (line color) (C). Text labels in panel C indicate the percent of area in an MPA at selected points along curves."
#| fig-height: 4
#| fig-width: 6

@@

tmp <- coral_results %>% 
  group_by(prop_mpa, placement_strategy) %>% 
  summarise(biodiv = sum(biodiv), yield = sum(yield)) %>% 
  group_by(placement_strategy) %>% 
  mutate(delta_yield = yield / yield[prop_mpa == 0] - 1,
         delta_biodiv = biodiv / biodiv[prop_mpa == 0] - 1) %>% 
  ungroup() %>% 
    mutate(placement_strategy= fct_relabel(placement_strategy, titler))

frontier_labels <- tmp %>% 
  mutate(delta_50 = (delta_yield - -0.5)^2,
         delta_0 = delta_yield^2) %>% 
  group_by(placement_strategy) %>% 
  filter(prop_mpa > 0) %>% 
  filter(yield == max(yield) | delta_0 == min(delta_0) | delta_50 == min(delta_50)) %>% 
  ungroup() %>% 
  mutate(pmpa = scales::percent(prop_mpa))

coral_frontier <- tmp %>% 
  ggplot(aes(delta_biodiv, delta_yield)) + 
  geom_hline(aes(yintercept = 0), linetype = 2) +
  geom_line(alpha = 0.8, aes(color = placement_strategy),size = 1.1) + 
  geom_text_repel(data = frontier_labels, aes(label = pmpa), min.segment.length = 0, box.padding = 0.5) +
  scale_x_continuous(name = "Change in Total Depletion", labels = scales::label_percent(accuracy = 1), guide = guide_axis(n.dodge = 2)) +
  scale_y_continuous(name = "Change in Yield", labels = scales::label_percent(accuracy = 1)) + 
  scale_color_manual(values = c("tomato", "steelblue"),name = '') +
  guides(fill = guide_legend(order = 1, override.aes = list(size = 4))) + 
  labs(tag = "C")  + 
  theme(
    legend.position = c(0.725, .95),
    legend.background = element_rect(fill = "transparent"),
    legend.text = element_text(size = 8)
  )

@@

```{r}
#| label: fig-blue-results
#| fig-cap: "Change in yield (A) and depletion (spawning biomass divided by unfished spawning biomass) (B) as a function of MPA size and design strategy by fishing fleet and species. Blue lines indicate impacts of MPAs under status qup habitat, red impacts under climate-driven range shift."
#| fig-height: 6
#| fig-width: 6
